---
title: "Clustering_Assignment"
author: "Kaylar Fullington"
date: 02/27/2022
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
cluster_df <- read.csv("data/clustering-data.csv")

library(ggplot2)

cluster_scatter <- ggplot(cluster_df, aes(x = x, y = y)) + geom_point()
cluster_scatter

```

```{r}
library(class)
library(factoextra)

cluster_kmeans_2 <- kmeans(scale(cluster_df), 2)

cluster_2_plot <- fviz_cluster(cluster_kmeans_2, data = cluster_df, geom = "point", ggtheme = theme_bw())
cluster_2_plot
```

```{r}
# clusterDist <- function(p1,p2) sqrt((p1[,1] - p2[,1])^2 + (p1[,2] - p2[,2]^2))
# cluster2_dist <- clusterDist(cluster_df[cluster_kmeans_2$cluster == 2], cluster_kmeans_2$centers[,2,drop = FALSE])
# Error in `[.data.frame`(cluster_df, cluster_kmeans_2$cluster == 2) : undefined columns selected

# distance <- map_dbl(2:12, function(k){ model <- kmeans(cluster_df, centers = k) model$distance})
# Error: unexpected symbol in "distance <- map_dbl(2:12, function(k){ model <- kmeans(cluster_df, centers = k) model"

distance_2 <- sqrt(rowSums(cluster_df - fitted(cluster_kmeans_2))^2)
av_dist_2 <- mean(distance_2)

```

```{r}
library(class)
library(factoextra)

cluster_kmeans_3 <- kmeans(scale(cluster_df), 3, nstart = 25)

cluster_3_plot <- fviz_cluster(cluster_kmeans_3, data = cluster_df, geom = "point", ggtheme = theme_bw())
cluster_3_plot
```

```{r}
distance_3 <- sqrt(rowSums(cluster_df - fitted(cluster_kmeans_3))^2)
av_dist_3 <- mean(distance_3)

```

```{r}
library(class)
library(factoextra)

cluster_kmeans_4 <- kmeans(scale(cluster_df), 4, nstart = 25)

cluster_4_plot <- fviz_cluster(cluster_kmeans_4, data = cluster_df, geom = "point", ggtheme = theme_bw())
cluster_4_plot
```

```{r}
distance_4 <- sqrt(rowSums(cluster_df - fitted(cluster_kmeans_4))^2)
av_dist_4 <- mean(distance_4)

```


```{r}
library(class)
library(factoextra)

cluster_kmeans_5 <- kmeans(scale(cluster_df), 5, nstart = 25)

cluster_5_plot <- fviz_cluster(cluster_kmeans_5, data = cluster_df, geom = "point", ggtheme = theme_bw())
cluster_5_plot
```

```{r}

distance_5 <- sqrt(rowSums(cluster_df - fitted(cluster_kmeans_5))^2)
av_dist_5 <- mean(distance_5)

```

```{r}
library(class)
library(factoextra)

cluster_kmeans_6 <- kmeans(scale(cluster_df), 6, nstart = 25)

cluster_6_plot <- fviz_cluster(cluster_kmeans_6, data = cluster_df, geom = "point", ggtheme = theme_bw())
cluster_6_plot
```

```{r}
distance_6 <- sqrt(rowSums(cluster_df - fitted(cluster_kmeans_6))^2)
av_dist_6 <- mean(distance_6)

```

```{r}
library(class)
library(factoextra)

cluster_kmeans_7 <- kmeans(scale(cluster_df), 7, nstart = 25)

cluster_7_plot <- fviz_cluster(cluster_kmeans_7, data = cluster_df, geom = "point", ggtheme = theme_bw())
cluster_7_plot
```

```{r}

distance_7 <- sqrt(rowSums(cluster_df - fitted(cluster_kmeans_7))^2)
av_dist_7 <- mean(distance_7)

```

```{r}
library(class)
library(factoextra)

cluster_kmeans_8 <- kmeans(scale(cluster_df), 8, nstart = 25)

cluster_8_plot <- fviz_cluster(cluster_kmeans_8, data = cluster_df, geom = "point", ggtheme = theme_bw())
cluster_8_plot
```

```{r}

distance_8 <- sqrt(rowSums(cluster_df - fitted(cluster_kmeans_8))^2)
av_dist_8 <- mean(distance_8)

```

```{r}
library(class)
library(factoextra)

cluster_kmeans_9 <- kmeans(scale(cluster_df), 9, nstart = 25)

cluster_9_plot <- fviz_cluster(cluster_kmeans_9, data = cluster_df, geom = "point", ggtheme = theme_bw())
cluster_9_plot
```

```{r}

distance_9 <- sqrt(rowSums(cluster_df - fitted(cluster_kmeans_9))^2)
av_dist_9 <- mean(distance_9)

```

```{r}
library(class)
library(factoextra)

cluster_kmeans_10 <- kmeans(scale(cluster_df), 10, nstart = 25)

cluster_10_plot <- fviz_cluster(cluster_kmeans_10, data = cluster_df, geom = "point", ggtheme = theme_bw())
cluster_10_plot
```

```{r}

distance_10 <- sqrt(rowSums(cluster_df - fitted(cluster_kmeans_10))^2)
av_dist_10 <- mean(distance_10)

```

```{r}
library(class)
library(factoextra)

cluster_kmeans_11 <- kmeans(scale(cluster_df), 11, nstart = 25)

cluster_11_plot <- fviz_cluster(cluster_kmeans_11, data = cluster_df, geom = "point", ggtheme = theme_bw())
cluster_11_plot
```

```{r}

distance_11 <- sqrt(rowSums(cluster_df - fitted(cluster_kmeans_11))^2)
av_dist_11 <- mean(distance_11)

```

```{r}
library(class)
library(factoextra)

cluster_kmeans_12 <- kmeans(scale(cluster_df), 12, nstart = 25)

cluster_12_plot <- fviz_cluster(cluster_kmeans_12, data = cluster_df, geom = "point", ggtheme = theme_bw())
cluster_12_plot
```

```{r}

distance_12 <- sqrt(rowSums(cluster_df - fitted(cluster_kmeans_12))^2)
av_dist_12 <- mean(distance_12)

```


```{r}
library(useful)
cluster_best <- FitKMeans(cluster_df, max.clusters = 12, nstart = 12)
cluster_best

PlotHartigan(cluster_best)

```


# Final Thoughts

I had to do a lot of messing around with this data and I finally hit a wall when I realized the average distances were all the same for every K-value. I am unsure what I did incorrectly that made this happen.

My clusters all seemed to come out right. At the bottom is Hartigan's method which in my understanding indicates similar to manually plotting out all these k-values. One can look at the bend to determine how many k-values would be best to use.


